1.4 Data Visualization with ggplot2 - R Practice

### 2019-09-06 content/r/03-r-practice.Rmd
Cosmetic tweaks to r practice 3, two #17s
Apply cosmetic change to r practice 3 page
Fix formatting for R practice 3 and 4, add answers for practice 3

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```
- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/r/03-r-practice.pdf)

@@

## 1.

a. Create a vector called `me` with two objects, your first name, and your last name. 
b. Call the vector to inspect it.
c. Confirm it is a character class vector.

## 2. Use R's help functions to determine what the `paste()` function does. Then paste together your first name and last name.

## 3. Create a vector called `my_vector` with all the even integers from 2 to 10.

## 4. Find the mean of `my_vector` with `mean()`.

## 5. Take all the integers from 18 to 763,^[Hint: use the `:` operator to create a sequence from a starting number to an ending number] then get the mean.

## Playing with Data

@@

### 6. Install `ggplot2`.

### 7. Load `ggplot2` with the `library()` command.

### 8. Get the `str`ucture of the diamonds data frame. What are the different variables and what kind of data does each contain?

### 9. Get summary statistics separately for `carat`, `depth`, `table`, and `price`.

### 10. `color`, `cut`, and `clarity` are categorical variables (factors). Use the `table()` command to generate frequency tables for each.

### 11. Now rerun the `summary()` command on the entire data frame.

### 12. Now look only at (subset) the first 4 diamonds in the dataset.

### 13. Now look only at (subset) the third and seventh diamond in the dataset.

### 14. Now look only at (subset) the second column of the dataset.

### 15. Do this again, but look using the `$` to pull up the second column by name.

### 16. Now look only at diamonds that have a `carat` greater than or equal to 1.

### 17. Now look only at diamonds that have a VVS1 `clarity`.

### 17. Now look only at dimaonds that have a color of E, F, I, and J.

### 18. Now look only at diamonds that have a `carat` greater than or equal to 1 and a VVS1 `clarity`.

### 19. Get the average price of diamonds in question 18.^[Hints: use your subset command as an argument to the mean function. You will not need a comma here because you are looking for a single row.]

### 20. What is the highest price for a diamond with a 1.0 carat, D color, and VVS1 clarity?

## Execute your R Script

>>>>

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
source(file.path(here::here(), "lib", "graphics.R"))
```
- [<i class="fas fa-file-pdf"></i> `Download as PDF`](/r/03-r-practice.pdf)
- [<i class="fas fa-file-pdf"></i> `Answer Key`](/r/03-r-practice-answers.pdf)
- [<i class="fas fa-registered"></i> `R Script`](/r/03-r-practice-answers.R)

@@

## 1

Work on the following parts:

- a. Create a vector called `me` with two objects, your first name, and your last name.
- b. Call the vector to inspect it.
- c. Confirm it is a character class vector.

## 2

Use R's help functions to determine what the `paste()` function does. Then paste together your first name and last name.

## 3

Create a vector called `my_vector` with all the even integers from 2 to 10.

## 4

Find the mean of `my_vector` with `mean()`.

## 5

Take all the integers from 18 to 763,^[Hint: use the `:` operator to create a sequence from a starting number to an ending number] then get the mean.

# Playing with Data

@@

## 6

Install `ggplot2`.

## 7

Load `ggplot2` with the `library()` command.

## 8

Get the `str`ucture of the diamonds data frame. What are the different variables and what kind of data does each contain?

## 9

Get summary statistics separately for `carat`, `depth`, `table`, and `price`.

## 10

`color`, `cut`, and `clarity` are categorical variables (factors). Use the `table()` command to generate frequency tables for each.

## 11

Now rerun the `summary()` command on the entire data frame.

## 12

Now look only at (subset) the first 4 diamonds in the dataset.

## 13

Now look only at (subset) the third and seventh diamond in the dataset.

## 14

Now look only at (subset) the second column of the dataset.

## 15

Do this again, but look using the `$` to pull up the second column by name.

## 16

Now look only at diamonds that have a `carat` greater than or equal to 1.

## 17

Now look only at diamonds that have a VVS1 `clarity`.

## 18

Now look only at dimaonds that have a color of E, F, I, and J.

## 19

Now look only at diamonds that have a `carat` greater than or equal to 1 and a VVS1 `clarity`.

## 20

Get the average price of diamonds in question 18.^[Hints: use your subset command as an argument to the mean function. You will not need a comma here because you are looking for a single row.]

## 21

What is the highest price for a diamond with a 1.0 carat, D color, and VVS1 clarity?

# Execute your R Script

### 2019-09-06 content/r/04-r-practice.Rmd
Cosmetic changes to r practice webpages for classes 3 and 4
Fix formatting for R practice 3 and 4, add answers for practice 3

## 1. We will look at GDP per Capita and Life Expectancy using some data from the [gapminder project](http://gapminder.org). There is a [handy package](https://github.com/jennybc/gapminder) called `gapminder` that uses a small snippet of this data for exploratory analysis. Install and load the package `gapminder`. Type `?gapminder` and hit enter to see a description of the data.

## 2. Let's get a quick look at `gapminder` to see what we're dealing with.

@@

## 3. Let's make sure you can do some basic plots before we get into the `gg`. Use base `R`'s `hist()` function to plot a *histogram* of `gdpPercap`.

## 4. Use base `R`'s `boxplot()` function to plot a *boxplot* of `gdpPercap`.

## 5. Now make it a *boxplot* by `continent`.^[Hint: use formula notation with`~`.]

## 6. Now make a *scatterplot* of `gdpPercap` on the $x$-axis and `LifeExp` on the $y$-axis.

@@

## 7. Load the package `ggplot2` (you should have installed it previously. If not, install first with `install.packages("ggplot2")`).

## 8. Let's first make a `bar` graph to see how many countries are in each continent. The only `aes`thetic you need is to map `continent` to `x`. Bar graphs are great for representing categories, but not quantitative data.

## 9. For quantitative data, we want a `histogram` to visualize the distribution of a variable. Make a `histogram` of `gdpPercap`. Your only `aes`thetic here is to map `gdpPercap` to `x`.

## 10. Now let's try adding some color, specifically, add an `aes`thetic that maps `continent` to `color.`

## 11. Instead of a `histogram`, change the `geom` to make it a `density` graph. To avoid overplotting, add `alpha=0.4` to the `geom` argument (alpha changes the *transparency* of a `fill`).

## 12. Redo your plot from 10 for `lifeExp` instead of `gdpPercap`.

## 13. Now let's try a scatterplot for `lifeExp` (as `y`) on `gdpPercap` (as `x`). You'll need both for `aes`thetics. The `geom` here is `geom_point()`.

## 14. Add some color by mapping `continent` to `color` in your `aes`thetics.

## 15. Now let's try adding a regression line with `geom_smooth()`. Add this layer on top of your `geom_point()` layer.

## 16. Did you notice that you got multiple regression lines (colored by continent)? That's because we set a `global` `aes`thetic of mapping `continent` to `color`. If we want just *one* regression line, we need to instead move the `color = continent` inside the `aes` of `geom_point`. This will only map `continent` to `color` for points, not for anything else.

## 17. Now add an `aes`thetic to your `point`s to map `pop` to `size`.

## 18. Change the color of the regression line to `"black"`. Try first by putting this inside an `aes()` in your `geom_smooth`, and try a second time by just putting it inside `geom_smooth` without an `aes()`. What's the difference, and why?

## 19. Another way to separate out continents is with `facet`ing. Add `+facet_wrap(~continent)` to create subplots by `continent`.

## 20. Remove the `facet` layer. The `scale` is quite annoying for the `x`-axis, a lot of points are clustered on the lower level. Let's try changing the scale by adding a layer: `+scale_x_log10()`.

## 21. Now let's fix the labels by adding `+labs()`. Inside `labs`, make proper axes titles for `x`, `y`, and a `title` to the plot. If you want to change the name of the legends (continent color), add one for `color` and `size`.

## 22. Now let's try subsetting by looking only at North America. Take the `gapminder` dataframe and subset it to only look at `continent=="Americas"`). Assign this to a new dataframe object (call it something like `america`.) Now, use *this* as your `data`, and redo the graph from question 16. (You might want to take a look at your new dataframe to make sure it worked first!)

## 23. Try this again for the *whole* world, but just for observations in the year 2000.

>>>>

## 1

We will look at GDP per Capita and Life Expectancy using some data from the [gapminder project](http://gapminder.org). There is a [handy package](https://github.com/jennybc/gapminder) called `gapminder` that uses a small snippet of this data for exploratory analysis. Install and load the package `gapminder`. Type `?gapminder` and hit enter to see a description of the data.

## 2

Let's get a quick look at `gapminder` to see what we're dealing with.

@@

## 3

Let's make sure you can do some basic plots before we get into the `gg`. Use base `R`'s `hist()` function to plot a *histogram* of `gdpPercap`.

## 4

Use base `R`'s `boxplot()` function to plot a *boxplot* of `gdpPercap`.

## 5

Now make it a *boxplot* by `continent`.^[Hint: use formula notation with `~`.]

## 6

Now make a *scatterplot* of `gdpPercap` on the $x$-axis and `LifeExp` on the $y$-axis.

@@

## 7

Load the package `ggplot2` (you should have installed it previously. If not, install first with `install.packages("ggplot2")`).

## 8

Let's first make a `bar` graph to see how many countries are in each continent. The only `aes`thetic you need is to map `continent` to `x`. Bar graphs are great for representing categories, but not quantitative data.

## 9

For quantitative data, we want a `histogram` to visualize the distribution of a variable. Make a `histogram` of `gdpPercap`. Your only `aes`thetic here is to map `gdpPercap` to `x`.

## 10

Now let's try adding some color, specifically, add an `aes`thetic that maps `continent` to `fill.`^[In general, `color` refers to the outside borders of a `geom` (except points), `fill` is the interior of an object.]

## 11

Instead of a `histogram`, change the `geom` to make it a `density` graph. To avoid overplotting, add `alpha=0.4` to the `geom` argument (alpha changes the *transparency* of a `fill`).

## 12

Redo your plot from 11 for `lifeExp` instead of `gdpPercap`.

## 13

Now let's try a scatterplot for `lifeExp` (as `y`) on `gdpPercap` (as `x`). You'll need both for `aes`thetics. The `geom` here is `geom_point()`.

## 14

Add some color by mapping `continent` to `color` in your `aes`thetics.

## 15

Now let's try adding a regression line with `geom_smooth()`. Add this layer on top of your `geom_point()` layer.

## 16

Did you notice that you got multiple regression lines (colored by continent)? That's because we set a `global` `aes`thetic of mapping `continent` to `color`. If we want just *one* regression line, we need to instead move the `color = continent` inside the `aes` of `geom_point`. This will only map `continent` to `color` for points, not for anything else.

## 17

Now add an `aes`thetic to your `point`s to map `pop` to `size`.

## 18

Change the color of the regression line to `"black"`. Try first by putting this inside an `aes()` in your `geom_smooth`, and try a second time by just putting it inside `geom_smooth` without an `aes()`. What's the difference, and why?

## 19

Another way to separate out continents is with `facet`ing. Add `+facet_wrap(~continent)` to create subplots by `continent`.

## 20

Remove the `facet` layer. The `scale` is quite annoying for the `x`-axis, a lot of points are clustered on the lower level. Let's try changing the scale by adding a layer: `+scale_x_log10()`.

## 21

Now let's fix the labels by adding `+labs()`. Inside `labs`, make proper axes titles for `x`, `y`, and a `title` to the plot. If you want to change the name of the legends (continent color), add one for `color` and `size`.

## 22

Now let's try subsetting by looking only at North America. Take the `gapminder` dataframe and subset it to only look at `continent=="Americas"`). Assign this to a new dataframe object (call it something like `america`.) Now, use *this* as your `data`, and redo the graph from question 17. (You might want to take a look at your new dataframe to make sure it worked first!)

## 23

Try this again for the *whole* world, but just for observations in the year 2002.

### 2019-09-06 static/r/04-r-practice.Rmd
Fix formatting for R practice 3 and 4, add answers for practice 3

## 1. We will look at GDP per Capita and Life Expectancy using some data from the [gapminder project](http://gapminder.org). There is a [handy package](https://github.com/jennybc/gapminder) called `gapminder` that uses a small snippet of this data for exploratory analysis. Install and load the package `gapminder`. Type `?gapminder` and hit enter to see a description of the data.

## 2. Let's get a quick look at `gapminder` to see what we're dealing with.

a. Get the `str`ucture of the `gapminder` data. 
b. What variables are there?
c. Look at the `head` of the dataset to get an idea of what the data looks like. 
d. Get `summary` statistics of all variables.

@@

## 3. Let's make sure you can do some basic plots before we get into the `gg`. Use base `R`'s `hist()` function to plot a *histogram* of `gdpPercap`.

## 4. Use base `R`'s `boxplot()` function to plot a *boxplot* of `gdpPercap`.

## 5. Now make it a *boxplot* by `continent`.^[Hint: use formula notation with`~`.]

## 6. Now make a *scatterplot* of `gdpPercap` on the $x$-axis and `LifeExp` on the $y$-axis.

@@

## 7. Load the package `ggplot2` (you should have installed it previously. If not, install first with `install.packages("ggplot2")`).

## 8. Let's first make a `bar` graph to see how many countries are in each continent. The only `aes`thetic you need is to map `continent` to `x`. Bar graphs are great for representing categories, but not quantitative data.

## 9. For quantitative data, we want a `histogram` to visualize the distribution of a variable. Make a `histogram` of `gdpPercap`. Your only `aes`thetic here is to map `gdpPercap` to `x`.

## 10. Now let's try adding some color, specifically, add an `aes`thetic that maps `continent` to `color.`

## 11. Instead of a `histogram`, change the `geom` to make it a `density` graph. To avoid overplotting, add `alpha=0.4` to the `geom` argument (alpha changes the *transparency* of a `fill`).

## 12. Redo your plot from 10 for `lifeExp` instead of `gdpPercap`.

## 13. Now let's try a scatterplot for `lifeExp` (as `y`) on `gdpPercap` (as `x`). You'll need both for `aes`thetics. The `geom` here is `geom_point()`.

## 14. Add some color by mapping `continent` to `color` in your `aes`thetics.

## 15. Now let's try adding a regression line with `geom_smooth()`. Add this layer on top of your `geom_point()` layer.

## 16. Did you notice that you got multiple regression lines (colored by continent)? That's because we set a `global` `aes`thetic of mapping `continent` to `color`. If we want just *one* regression line, we need to instead move the `color = continent` inside the `aes` of `geom_point`. This will only map `continent` to `color` for points, not for anything else.

## 17. Now add an `aes`thetic to your `point`s to map `pop` to `size`.

## 18. Change the color of the regression line to `"black"`. Try first by putting this inside an `aes()` in your `geom_smooth`, and try a second time by just putting it inside `geom_smooth` without an `aes()`. What's the difference, and why?

## 19. Another way to separate out continents is with `facet`ing. Add `+facet_wrap(~continent)` to create subplots by `continent`.

## 20. Remove the `facet` layer. The `scale` is quite annoying for the `x`-axis, a lot of points are clustered on the lower level. Let's try changing the scale by adding a layer: `+scale_x_log10()`.

## 21. Now let's fix the labels by adding `+labs()`. Inside `labs`, make proper axes titles for `x`, `y`, and a `title` to the plot. If you want to change the name of the legends (continent color), add one for `color` and `size`.

## 22. Now let's try subsetting by looking only at North America. Take the `gapminder` dataframe and subset it to only look at `continent=="Americas"`). Assign this to a new dataframe object (call it something like `america`.) Now, use *this* as your `data`, and redo the graph from question 16. (You might want to take a look at your new dataframe to make sure it worked first!)

## 23. Try this again for the *whole* world, but just for observations in the year 2000.

>>>>

### 1. We will look at GDP per Capita and Life Expectancy using some data from the [gapminder project](http://gapminder.org). There is a [handy package](https://github.com/jennybc/gapminder) called `gapminder` that uses a small snippet of this data for exploratory analysis. Install and load the package `gapminder`. Type `?gapminder` and hit enter to see a description of the data.

### 2. Let's get a quick look at `gapminder` to see what we're dealing with.

a. Get the `str`ucture of the `gapminder` data.

b. What variables are there?

c. Look at the `head` of the dataset to get an idea of what the data looks like.

d. Get `summary` statistics of all variables.

@@

### 3. Let's make sure you can do some basic plots before we get into the `gg`. Use base `R`'s `hist()` function to plot a *histogram* of `gdpPercap`.

### 4. Use base `R`'s `boxplot()` function to plot a *boxplot* of `gdpPercap`.

### 5. Now make it a *boxplot* by `continent`.^[Hint: use formula notation with `~`.]

### 6. Now make a *scatterplot* of `gdpPercap` on the $x$-axis and `LifeExp` on the $y$-axis.

@@

### 7. Load the package `ggplot2` (you should have installed it previously. If not, install first with `install.packages("ggplot2")`).

### 8. Let's first make a `bar` graph to see how many countries are in each continent. The only `aes`thetic you need is to map `continent` to `x`. Bar graphs are great for representing categories, but not quantitative data.

### 9. For quantitative data, we want a `histogram` to visualize the distribution of a variable. Make a `histogram` of `gdpPercap`. Your only `aes`thetic here is to map `gdpPercap` to `x`.

### 10. Now let's try adding some color, specifically, add an `aes`thetic that maps `continent` to `fill`.^[In general, `color` refers to the outside borders of a `geom` (except points), `fill` is the interior of an object.]

### 11. Instead of a `histogram`, change the `geom` to make it a `density` graph. To avoid overplotting, add `alpha=0.4` to the `geom` argument (alpha changes the *transparency* of a `fill`).

### 12. Redo your plot from 11 for `lifeExp` instead of `gdpPercap`.

### 13. Now let's try a scatterplot for `lifeExp` (as `y`) on `gdpPercap` (as `x`). You'll need both for `aes`thetics. The `geom` here is `geom_point()`.

### 14. Add some color by mapping `continent` to `color` in your `aes`thetics.

### 15. Now let's try adding a regression line with `geom_smooth()`. Add this layer on top of your `geom_point()` layer.

### 16. Did you notice that you got multiple regression lines (colored by continent)? That's because we set a `global` `aes`thetic of mapping `continent` to `color`. If we want just *one* regression line, we need to instead move the `color = continent` inside the `aes` of `geom_point`. This will only map `continent` to `color` for points, not for anything else.

### 17. Now add an `aes`thetic to your `point`s to map `pop` to `size`.

### 18. Change the color of the regression line to `"black"`. Try first by putting this inside an `aes()` in your `geom_smooth`, and try a second time by just putting it inside `geom_smooth` without an `aes()`. What's the difference, and why?

### 19. Another way to separate out continents is with `facet`ing. Add `+facet_wrap(~continent)` to create subplots by `continent`.

### 20. Remove the `facet` layer. The `scale` is quite annoying for the `x`-axis, a lot of points are clustered on the lower level. Let's try changing the scale by adding a layer: `+scale_x_log10()`.

### 21. Now let's fix the labels by adding `+labs()`. Inside `labs`, make proper axes titles for `x`, `y`, and a `title` to the plot. If you want to change the name of the legends (continent color), add one for `color` and `size`.

### 22. Now let's try subsetting by looking only at North America. Take the `gapminder` dataframe and subset it to only look at `continent=="Americas"`). Assign this to a new dataframe object (call it something like `america`.) Now, use *this* as your `data`, and redo the graph from question 17. (You might want to take a look at your new dataframe to make sure it worked first!)

### 23. Try this again for the *whole* world, but just for observations in the year 2002.
